---
title: "PM566 WEBSITE FINAL"
author: "Erin Yu"
embed-resources: true
echo: FALSE
message: false
warning: false
---
This final project utilized the 2015-2022 Alzheimer's Disease and Healthy Aging Data from the CDC Division of Population Health's Behavioral Risk Factor Surveillance System (BRFSS) to evaluate mental health struggles in people with Alzheimer's Disease. The results showed that individuals in the Southern states reported the highest percentages of frequent mental distress and lifetime diagnosis of depression. Of all the racial/ethnic groups, Native American/Alaskan Natives reported the highest rates of frequent mental distress (18.5%) and lifetime diagnosis of depression (25.1%). 

You can download the PDF version of the report [here](PM-566-FINAL.pdf).


```{r}
# Read the CSV file into a data frame
data <- read.csv("/Users/erinyu/Desktop/USC/MPH/Fall 2024/Health Data (PM 566)/Alzheimer_s_Disease_and_Healthy_Aging_Data.csv")

```


```{r}
# Create a subset of the data where Class is "Mental Health"

mental_health_data <- subset(data, Class == "Mental Health")


```



```{r}
# Remove rows with NA in Data_Value

clean_mentalhealth_data <- mental_health_data[!is.na(mental_health_data$Data_Value), ]


```
This is an interactive table reporting the average percentage reporting frequent mental distress by each state. 

```{r}
# Load necessary libraries
library(dplyr)
library(DT)

# Filter for "Frequent mental distress" under Topic and calculate the mean percentage
summarized_data <- clean_mentalhealth_data %>%
  filter(Topic == "Frequent mental distress") %>%
  group_by(LocationDesc) %>%
  summarize(
    `Mean percentage` = mean(Data_Value, na.rm = TRUE)
  ) %>%
  rename(State = LocationDesc)  # Rename column header to "State"

# Create an interactive table with a custom title
datatable(
  summarized_data,
  options = list(
    pageLength = 10,  # Show 10 rows per page
    autoWidth = TRUE  # Adjust column widths
  ),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: center; font-size: 16px; font-weight: bold;",
    "Mean Percentage Reporting Frequent Mental Distress by State"
  )
)


```
This is a heat map of the mean percentages reporting frequent mental distress by state. 
```{r}
library(leaflet)
library(dplyr)
library(tigris)
library(sf)
options(tigris_use_cache = TRUE)

# Filter and calculate mean percentages
summarized_data <- clean_mentalhealth_data %>%
  filter(Topic == "Frequent mental distress") %>%
  group_by(LocationDesc) %>%
  summarize(Mean_Percentage = mean(Data_Value, na.rm = TRUE)) %>%
  rename(State = LocationDesc)  # Rename for easier matching

# Load state boundaries (simplified for efficiency)
states <- tigris::states(cb = TRUE, class = "sf")

# Ensure state names match by converting to uppercase
summarized_data <- summarized_data %>%
  mutate(State = toupper(State))

# Join state shapefile with summarized data
states_data <- states %>%
  mutate(NAME = toupper(NAME)) %>%  # Ensure NAME matches State
  left_join(summarized_data, by = c("NAME" = "State"))

# Create a Leaflet map and add a centered title
leaflet(data = states_data) %>%
  addPolygons(
    fillColor = ~colorQuantile("YlOrRd", Mean_Percentage)(Mean_Percentage),
    weight = 1,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 2,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste(NAME, "<br>", "Mean %: ", round(Mean_Percentage, 2))
  ) %>%
  addLegend(
    "bottomright",
    pal = colorQuantile("YlOrRd", domain = states_data$Mean_Percentage, n = 5),
    values = ~Mean_Percentage,
    title = "Mean % Reporting Frequent Mental Distress",
    opacity = 1
  ) %>%
  # Add title as a top-left control with custom HTML and CSS
  addControl(
    html = "<div style='text-align:center; font-size:16px; font-weight:bold;'>Mean Percentage Reporting Frequent Mental Distress by State</div>",
    position = "topleft"
  )

```
This is an interactive table reporting the average percentages of individuals reporting a lifetime diagnosis of depression by state

```{r}
# Load necessary libraries
library(dplyr)
library(DT)

# Filter for "Lifetime diagnosis of depression" under Topic and calculate the mean percentage
summarized_depression_data <- clean_mentalhealth_data %>%
  filter(Topic == "Lifetime diagnosis of depression") %>%
  group_by(LocationDesc) %>%
  summarize(
    `Mean percentage` = mean(Data_Value, na.rm = TRUE)
  ) %>%
  rename(State = LocationDesc)  # Rename column header to "State"

# Create an interactive table with a custom title
datatable(
  summarized_depression_data,
  options = list(
    pageLength = 10,  # Show 10 rows per page
    autoWidth = TRUE  # Adjust column widths
  ),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: center; font-size: 16px; font-weight: bold;",
    "Mean Percentage Reporting Lifetime Diagnosis of Depression by State"
  )
)

```
This is a heat map of the mean percentages reporting lifetime diagnosis of depression by state. 
```{r}
library(leaflet)
library(dplyr)
library(tigris)
library(sf)
options(tigris_use_cache = TRUE)

# Filter and calculate mean percentages for Lifetime Diagnosis of Depression
summarized_data <- clean_mentalhealth_data %>%
  filter(Topic == "Lifetime diagnosis of depression") %>%
  group_by(LocationDesc) %>%
  summarize(Mean_Percentage = mean(Data_Value, na.rm = TRUE)) %>%
  rename(State = LocationDesc)  # Rename for easier matching

# Load state boundaries (simplified for efficiency)
states <- tigris::states(cb = TRUE, class = "sf")

# Ensure state names match by converting to uppercase
summarized_data <- summarized_data %>%
  mutate(State = toupper(State))

# Join state shapefile with summarized data
states_data <- states %>%
  mutate(NAME = toupper(NAME)) %>%  # Ensure NAME matches State
  left_join(summarized_data, by = c("NAME" = "State"))

# Create a Leaflet map for Lifetime Diagnosis of Depression
leaflet(data = states_data) %>%
  addPolygons(
    fillColor = ~colorQuantile("YlOrRd", Mean_Percentage)(Mean_Percentage),
    weight = 1,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 2,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste(NAME, "<br>", "Mean %: ", round(Mean_Percentage, 2))
  ) %>%
  addLegend(
    "bottomright",
    pal = colorQuantile("YlOrRd", domain = states_data$Mean_Percentage, n = 5),
    values = ~Mean_Percentage,
    title = "Mean % Diagnosed with Depression",
    opacity = 1
  ) %>%
  # Add title as a top-left control with custom HTML and CSS
  addControl(
    html = "<div style='text-align:center; font-size:16px; font-weight:bold;'>Mean Percentage Diagnosed with Depression by State</div>",
    position = "topleft"
  )
```

